Bandit Multiclass Classification

### 2018-10-18 main.tex
ova algorithm

$M$ counts the number of different $(S_1,\ldots, S_\ell)$'s with $S_1\times \ldots \times S_\ell=S$. Since $S$ has degree at most $\deg$, we can bound $M$ by $ \ell^{\deg}$.

>>>>

$M$ counts the number of different $(S_1,\ldots, S_\ell)$'s with $S_1\times \ldots \times S_\ell=S$. Since $S$ has degree at most $\deg$, we can bound $M$ by $ \ell^{\deg}$. \footnote{This can be seen in two ways. First, we can think of ``assigning'' elements in $S$ to one of $(S_1,\ldots, S_\ell)$; every partition $S_1,\ldots,S_\ell$ can be achieved by some assignment - this can have at most $\ell^{\deg}$ ways. Second, consider $S = (s_1, \ldots, s_d)$.
Consider decomposing each coordinate of $S$ - coordinate $i$ has ${s_i + \ell - 1 \choose \ell - 1}$ ways to partition to $\ell$ parts. Therefore the total number of possible partitions is $\prod_{i=1}^d {s_i + \ell - 1 \choose \ell - 1} \leq \prod_{i=1}^d \ell^{s_i} \leq \ell^{\deg}$.}

### 2018-10-18 ptf_margin.tex
ova algorithm

\DeclareMathOperator*{\arginf}{arginf}
\DeclareMathOperator*{\argsup}{argsup}
\DeclareMathOperator*{\range}{range}
\DeclareMathOperator*{\mydet}{det_{+}}
%\DeclarePairedDelimiter\abs{\lvert}{\rvert}
\DeclarePairedDelimiter\bigabs{\big\lvert}{\big\rvert}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
\DeclarePairedDelimiter\bigceil{\big\lceil}{\big\rceil}
\DeclarePairedDelimiter\bigfloor{\big\lfloor}{\big\rfloor}

@@

\section{Learning intersection of halfspaces with a margin}
Suppose we are given examples $S = \cbr{(x, y)}$, where $x$ is in $\cbr{x \in \fR^d: \| x \| \leq 1}$, and $y$ is in $\cbr{-1,+1}$.
\begin{definition}
Dataset $S$ is said to be separable by an intersection of halfspaces with margin parameter $\gamma > 0$, if there exists $t$ linear classifiers $v_1, \ldots, v_t$ (all of which are in $\cbr{v \in \fR^d: \| v \| \leq 1}$),
such that for every $(x,y)$ in $S$:
\begin{enumerate}
  \item If $y = +1$, then $v_i \cdot x \geq \gamma$ holds simultaneously for all $i$ in $[t]$.
  \item If $y = -1$, then there exists an $i$ in $[t]$, $v_i \cdot x \leq -\gamma$.
\end{enumerate}
\label{def:int-margin}
\end{definition}
\cz{Looking at~\cite{klivans2004learning} and its proof, it assumes that for all $(x,y)$ in $S$, $|v_i \cdot x| \geq \gamma$. In our subsequent proofs, we show that it can be in fact weakened to the condition above, which is also more suitable for online bandit multiclass linear classification.}

@@

\section{Construction of large margin feature mappings using polynomials}
Our $\phi$'s in considerations are as follows:
\begin{enumerate}
%\sqrt{\frac{1}{\deg+1}}
\item $\phi_{s}(x) = ( x_1^{i_1} \ldots x_d^{i_d})_{i_1 + \ldots + i_d \leq s, i_1, \ldots, i_d \geq 0} = (x_S)_{S \in \fN^d, \|S\|_1 \leq s}$. This corresponds to the complete symmetric kernel $K_s$. Note that for all $x$ such that $\| x \|_2 \leq 1$,
$\inner{\phi_{s}(x), \phi_{s}(x)} \leq \sum_{i=0}^s \|x\|^{2i} \leq 1 + s$.

>>>>

\DeclareMathOperator*{\sep}{sep}
\DeclareMathOperator*{\arginf}{arginf}
\DeclareMathOperator*{\argsup}{argsup}
\DeclareMathOperator*{\range}{range}
\DeclareMathOperator*{\mydet}{det_{+}}
%\DeclarePairedDelimiter\abs{\lvert}{\rvert}
\DeclarePairedDelimiter\bigabs{\big\lvert}{\big\rvert}
\DeclarePairedDelimiter\ceil{\lceil}{\rceil}
\DeclarePairedDelimiter\floor{\lfloor}{\rfloor}
\DeclarePairedDelimiter\bigceil{\big\lceil}{\big\rceil}
\DeclarePairedDelimiter\bigfloor{\big\lfloor}{\big\rfloor}

@@

\section{Preliminaries}
\begin{definition}
$S \subset (\calX \times [K])^*$ is said to be {\em linearly separable} with margin parameter $\gamma > 0$, if there exists an multiclass linear classifier $W \in \fR^{K \times d}$ ($\| W \|_F \leq 1$),
such that for every $(x,y)$ in $S$,
$W_y \cdot x \geq W_{y'} \cdot x + \gamma \max_{(x,y) \in S} \| x \|_2$ holds simultaneously for all $y' \neq y$.
\label{def:mc-margin}
\end{definition}

\begin{definition}
$S \subset (\calX \times [K])^*$ is said to be {\em one-versus-all (OvA) linearly separable} with margin parameter $\gamma > 0$, if there exists an multiclass linear classifier $W \in \fR^{k \times d}$ ($\| W \|_F \leq 1$),
such that for every $(x,y)$ in $S$,
$W_y \cdot x \geq \gamma \max_{(x,y) \in S} \| x \|_2 $, 
and 
$W_{y'} \cdot x \leq -\gamma \max_{(x,y) \in S} \| x \|_2$
holds simultaneously for all $y' \neq y$.
\label{def:mc-ova-margin}
\end{definition}

For every example $(x, y)$, denote by 
\[ b_y^i = \begin{cases} +1  & y_t = i \\ -1 & y_t \neq i \end{cases} \]
its one-versus-all label. The above definition is equivalent to: there exists a $W$ in $\fR^{K \times d}$ ($\|W\|_F \leq 1$),
for every $i$ in $[K]$,
for every $(x, y)$ in $S$,
\[ 
b_y^i W_i \cdot x \geq \gamma \max_{(x,y) \in S} \| x \|_2.
\]

\cz{The reason we are using this specific notion of linear separability is:
1. We need a direct comparison of $\gamma$'s before and after feature transformation, therefore a ``normalized'' notion of margin is used.  
2. Alternative notions such as $\frac{W_y \cdot x}{\| x \|_2} \geq \gamma$ requires normalizing each example, in order to get a $\frac{1}{\gamma^2}$ mistake bound. If kernel Perceptron is used, this extra normalization step is equivalent to using a new kernel $k'(x,x') = \frac{k(x,x')}{\sqrt{k(x,x) k(x',x')}}$ as opposed to using the original kernel $k$ - we would like to avoid this techinicality for ease of presentation.}

\section{Main algorithm}
Our key observation is that, we can reduce the problem of online bandit linear multiclass learning under linear separability to the same problem under OvA linear separability. Specifically, inspired by~\cite{klivans2004learning}, we define
a feature map $\phi: \fR^d \to \fR^{d'}$, such that if the original dataset $S$ is linearly separable with margin $\gamma > 0$, then the dataset tranformed by $\phi$, $S' := \cbr{(\phi(x), y): (x,y) \in S}$ is OvA linearly separable with margin $\gamma' > 0$. Built on this feature map, we propose an efficient algorithm that has a finite mistake bound in the linearly separable setting. Its key idea is to 
run an algorithm that has a finite mistake bound under OvA linear separability, upon the transformed examples $(\phi(x), y)$. 
%a Perceptron-based algorithm that achieves a finite mistake bound under strong linear separablilty, and show how it can be combined with the feature map $\phi$ to get a
%online learning intersection of halfspaces.

\begin{algorithm}[H]
\caption{Main Algorithm}
\label{alg:ova}
\SetKwInOut{Input}{Input} 
\Input{Feature mapping $\phi$}
Initialize $w_t^{(1)}=\ldots=w_t^{(K)}=\mathbf{0}\in \mathbb{R}^{d'}$\\
\For{$t=1$ to $T$}{
   \If{$\exists y$ such that $w_t^{(y)\top} \phi(x_t) \geq 0$}{
       Assign $\tilde{y}_t$ to any $y$ with $w_t^{(y)\top}  \phi(x_t) \geq 0$ \\
       Predict $\tilde{y}_t$ \\
       \lIf{$\tilde{y}_t\neq y_t$}{ update $w_t^{(\tilde{y}_t)} \leftarrow w_t^{(\tilde{y}_t)} - \phi(x_t)$ } \label{line:update1}
   }
   \Else{
       Pick $\tilde{y}_t$ randomly from $
       \text{unif}([K])$ \label{line:explore}\\
       Predict $\tilde{y}_t$ \\
       \lIf{$\tilde{y}_t = y_t$}{ update $w_t^{(\tilde{y}_t)} \leftarrow w_t^{(\tilde{y}_t)} + \phi(x_t)$ } \label{line:update2}
   }
}
\end{algorithm}

We have the following guarantees of the above algorithm.
\begin{theorem}
Suppose feature map $\phi$ is such that for \cbr{(\phi(x_t), y_t): t \in [T]} is OvA linearly separable by margin $\gamma'$ except for $N$ points. Then, 
the expected cumulative number of mistakes made by Algorithm~\ref{alg:ova} is at most $K \cdot ( (1 + \frac 1 {\gamma'}) N + \sqrt{ (1 + \frac 1 {\gamma'}) N \cdot \frac{1}{\gamma'^2} } + \frac{1}{\gamma'^2})$.
\end{theorem}

\begin{corollary}
Suppose feature map $\phi$ is such that \cbr{(\phi(x_t), y_t): t \in [T]} is OvA linearly separable by margin $\gamma'$, then the cumulative number of mistakes made by Algorithm~\ref{alg:ova} is $\frac{K}{\gamma'^2}$.
\end{corollary}

\begin{proof}
Denote by $U_{t,1}$ (resp. $U_{t,2}$) the indicator in the ``if'' branch (resp. ``else'' branch) at time $t$. In addition, denote by $U_t^y$ the indicator of $w_t^{(y)}$ gets updated at 
round $t$. It can be easily seen that $U_{t,1} + U_{t,2} = \sum_{y=1}^k U_t^y = U_t^{\tilde{y}_t}$, which is equal to the indicator of Algorithm~\ref{alg:ova} making an update at iteration $t$. Let $N_{t,2}$ be the indicator of Algorithm~\ref{alg:ova} encounters the ``else'' branch at time $t$. Throughout, we also use the shorthand that $b_{t,i} = b_{y_t}^i$.

% for every $y$ in $[K]$, we have the following bound on $\sum_{t=1}^T U_t^y$: for all $w_y \in \fR^d$ and all $\rho > 0$,
%\begin{eqnarray*}
%\sum_{t=1}^T U_t^y 
%\leq
%\sum_{t=1}^T U_t^y (1 - \frac{b_{t,y} w_y \phi(x_t)}{\rho})_+ 
%+ \sqrt{(\sum_{t=1}^T U_t^y (1 - \frac{b_{t,y} w_y \phi(x_t)}{\rho})_+) \cdot \frac{\|w_y\|^2 R^2}{\rho^2} } 
%+ \frac{\|w_y\|^2 R^2}{\rho^2}
%\end{eqnarray*}
%
%Summing over all $y$ in $[K]$ and doing some algebra,

By standard analysis of the Perceptron algorithm, we have that for all $W \in \fR^{K \times d}$, and all $\rho > 0$,
\begin{equation}
\sum_{t=1}^T U_t^{\tilde{y}_t} 
\leq 
\sum_{t=1}^T  U_t^{\tilde{y}_t} (1 - \frac{b_{t,y} w_{\tilde{y}_t} \phi(x_t)}{\rho})_+ 
+ \sqrt{(\sum_{t=1}^T  U_t^{\tilde{y}_t} (1 - \frac{b_{t,y} w_{\tilde{y}_t} \phi(x_t)}{\rho})_+) \cdot \frac{\|W\|_F^2 R^2}{\rho^2} } 
+ \frac{\|W\|_F^2 R^2}{\rho^2},
\label{eqn:perceptron-mb}
\end{equation}
where $R = \max_{t=1}^T \| \phi(x_t)\|$.

Denote by $S_{\sep}$ the indices of examples in $\cbr{(x_t, y_t)}_{t=1}^T$ that is separable by a margin $\gamma$.
Take $W = W^*$, and $\rho = \gamma' \max_{t=1}^T \| \phi(x_t) \|_2$, we have that for all $t$ in $S_{\sep}$, for all $y$ in $[K]$, $(1 - \frac{b_{t,y} W^*_y \phi(x_t)}{\rho})_+ = 0$. For $t$ not in $S_{\sep}$,
$\sum_{y=1}^K U_t^y (1 - \frac{b_{t,y} W^*_y \phi(x_t)}{\rho})_+ \leq (1 + \frac{1}{\gamma'})$. 
This implies that $\sum_{t=1}^T  \sum_{y=1}^K U_t^y (1 - \frac{b_{t,y} w_y \phi(x_t)}{\rho})_+ \leq N(1 + \frac 1 {\gamma'})$.

Combining with Equation~\eqref{eqn:perceptron-mb}, we get 
\[
\sum_{t=1}^T U_t^{\tilde{y}_t}  \leq (1 + \frac 1 {\gamma'}) N + \sqrt{ (1 + \frac 1 {\gamma'}) N \cdot \frac{1}{\gamma'^2} } + \frac{1}{\gamma'^2}.
\]

Now, observe that $M_t$, the mistake indicator at time $t$ is at most $U_{t,1} + N_{t,2}$. Also, note that $\E[U_{t,2} | N_{t,2}] \geq \frac 1 K \cdot N_{t,2}$, as conditioned on $N_{t,2} = 1$, with probability $\frac 1 K$, $U_{t,2} = 1$ . This implies that
\[
\E[\sum_{t=1}^T M_t ] \leq \E[\sum_{t=1}^T U_{t,1} + N_{t,2} ] \leq K \E[\sum_{t=1}^T U_{t,1} + U_{t,2} ]  \leq K ((1 + \frac 1 {\gamma'}) N + \sqrt{ (1 + \frac 1 {\gamma'}) N \cdot \frac{1}{\gamma'^2} } + \frac{1}{\gamma'^2}).
\]
\end{proof}

%\section{Learning intersection of halfspaces with a margin}

\section{Feature mappings that induce OvA linear separability from linear separability}

How do we come up with a feature map that maps a linearly separable dataset to a OvA linearly separable dataset? Recall Definition~\ref{def:mc-ova-margin}, it suffices to give a feature map $\phi$ that for every $i$ in $K$, dataset $S_i = \cbr{(x, b_y^i): (x,y) \in S}$ is separable after the transformation. What property does dataset $S^i$ possess? Observe that for all examples $(x,y)$ in $S$ such that $b_y^i = +1$ (i.e. $y = i$), $x$ is such that $(w_i - w_{y'}) \cdot x \geq \gamma$ for all $y' \neq i$. For all examples $(x,y)$ in $S$ such that $b_y^i = -1$ (i.e. $y \neq i$), $x$ is such that 
$(w_y - w_i) \cdot x \geq \gamma$, specifically, $(w_i - w_{y'}) \cdot x \leq -\gamma$ for some $y' \neq i$. This motivates the following definition.

%Suppose we are given examples 
\begin{definition}
Suppose we are given dataset $S = \cbr{(x, y)}$, where $x$ is in $\cbr{x \in \fR^d: \| x \| \leq 1}$, and $y$ is in $\cbr{-1,+1}$.
$S$ is said to be separable by an intersection of halfspaces with margin parameter $\gamma > 0$, if there exists $t$ linear classifiers $v_1, \ldots, v_t$ (all of which are in $\cbr{v \in \fR^d: \| v \| \leq 1}$),
such that for every $(x,y)$ in $S$:
\begin{enumerate}
  \item If $y = +1$, then $v_i \cdot x \geq \gamma$ holds simultaneously for all $i$ in $[t]$.
  \item If $y = -1$, then there exists an $i$ in $[t]$, $v_i \cdot x \leq -\gamma$.
\end{enumerate}
\label{def:int-margin}
\end{definition}
\cz{Looking at~\cite{klivans2004learning} and its proof, it assumes that for all $(x,y)$ in $S$, $|v_i \cdot x| \geq \gamma$. In our subsequent proofs, we show that it can be in fact weakened to the condition above, which is also more suitable for online bandit multiclass linear classification.}

@@

Our $\phi$'s in considerations are as follows:
\begin{enumerate}
%\sqrt{\frac{1}{\deg+1}}
\item $\phi_{s}(x) = ( x_1^{i_1} \ldots x_d^{i_d})_{i_1 + \ldots + i_d \leq s, i_1, \ldots, i_d \geq 0} = (x_S)_{S \in \fN^d, \|S\|_1 \leq s}$. This corresponds to the complete symmetric kernel $K_s$. Note that for all $x$ such that $\| x \|_2 \leq 1$,
$\inner{\phi_{s}(x), \phi_{s}(x)} \leq \sum_{i=0}^s \|x\|^{2i} \leq 1 + s$.
